---
title: "EDLD652 Final"
author: "Diana Dewald, Eliott Doyle"
date: "2/10/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(rio)
library(here)
library(tidytext)
library(RColorBrewer)

#opening and removing about 90% of the rows since the file was too big to push! This code won't work since I got rid of the blogtext.csv file 
#data1 <- import(here("blogtext.csv"))
#data1 <- data1[-(60000:681284),]
#write.csv(data1, "blogtextabbr.csv")

#reading in the new df with fewer observations
data <- import(here("blogtextabbr.csv"))
```

```{r setup2, include=FALSE}
#making all text in 'text' lowercase
data <- data %>% 
  mutate(text = tolower(text))

#View(data)

m_age <- mean(data$age, na.rm = T)
m_age
max_age <- max(data$age, na.rm = T)
max_age
min_age <- min(data$age, na.rm = T)
min_age

#creating a new column with three groups for age
data <- data %>% 
  mutate(agegroup = cut(age, 
                        breaks =c (-Inf, 19.9, 30, Inf), 
                        labels=c("Under 20","20s","Over 30")))
head(data)
```

# Description of the data
#### The data used for this project are from https://www.kaggle.com/rtatman/blog-authorship-corpus, a text dataset of approximately 681000 blogposts. For the purposes of this project, we will be working with one tenth of the full dataset. In addition to the text of each blog post, some information about each the post (topic, date posted) and blogger (gender, age, astrological sign) is included.

# Research question 1:
## What are the most frequently used words in all writing samples?

```{r rq1, warning= FALSE, echo = FALSE, message = FALSE}
#parsing words from the 'text' column
tidy_words <- data %>% 
  unnest_tokens(word, text)

#removing numbers
tidy_words <- tidy_words[-grep("\\b\\d+\\b", tidy_words$word),]

#removing common/under-informative words
exclu <- tibble(word = c("nbsp","urllink", "lot", "â", "im", "pm", "blog", "stuff", "lol"))
tidy_words <- tidy_words %>% 
  anti_join(exclu, by = "word")
head(tidy_words)

#counting most frequently occurring words, removing stop words from counts
# tidy_words %>% 
#  anti_join(stop_words) %>%
#  count(word, sort = TRUE)

#plot
#NOTE: Cool plot! I added in a geom_text code on line 87 where you could add the numbers to each word to show how frequent they are (side note, couldn't get the nudge to work so you'd have to play with that but it could be a cool addition)
tidy_words %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% # make y-axis ordered by n
  slice(1:15) %>% # select only the first 15 rows
  ggplot(aes(n, word)) +
  geom_col(fill = "cornflowerblue") +
  geom_text(
    aes(n, word, label = round(n, 2)),
    nudge_x = -2,
    size = 4
  ) +
  scale_x_continuous(expand = c(0,0)) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words across all blog posts",
    caption = "Data abbreviated from https://www.kaggle.com/rtatman/blog-authorship-corpus"
  )
```

# Research questions 2a: 
## What are the most frequently used words by age group? Which words are unique to certain age groups?

### For Final:
#### -color columns by words that are unique to each age group, as well as words that are used in only 2 age groups

```{r rq2a, warning= FALSE, echo = FALSE, message = FALSE}
# separate words by valence (positive and negative)?
# separate words by categories (time, place, noun, feeling)?


# what I tried initially to organize dataframe by age (wasn't ideal for plotting)
# words_age <- tidy_words %>%
#  mutate(under_20 = ifelse(agegroup == "Under 20", 1, 0),
#         "20s" = ifelse(agegroup == "20s", 1, 0),
#         over_30 = ifelse(agegroup == "Over 30", 1, 0)) %>%
#  select(under_20, word)

#also tried group_by(agegroup) in initial plot, this didn't work, nor did facet_wrap(~agegroup)
#I know the code below is not the most elegant way to group by age, so I'm open to suggestions here

# note: I attempted to color bars by uniqueness (see 'for final' not above), but leftjoin() was giving me trouble


# prepping data
tidy <- tidy_words %>% 
  group_by(agegroup) %>%
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% 
  slice(1:15)

age <- tidy %>%
  distinct() %>%
  pivot_wider(names_from = agegroup, values_from = word) 


# plotting younget age group
n_under20 <- age$n[1:15]
under_20 <- age$`Under 20`[1:15]

young <- data.frame(n_under20, under_20)

#NOTE: again thought it would be cool to leave in the number to the plot, maybe throughout if you have the same plots! Also thought it would be cool if you added in a ggtext to your title to color some words to match the col color. I added in some code below you could use for that. 
  ggplot(young, aes(n_under20, under_20)) +
  geom_col(fill = "lightpink4") +
    geom_text(
    aes(n_under20, under_20, label = round(n_under20, 2)),
    nudge_x = -2,
    size = 4
  ) +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those <span style = 'color: lightpink4'>**Under 20**</span>",
  ) +
    theme(plot.title = ggtext::element_markdown()) 


# plotting middle age group--cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
n_20s <- age$n[16:30]
twenties <- age$`20s`[16:30]

middle <- data.frame(n_20s, twenties)

  ggplot(middle, aes(n_20s, twenties)) +
  geom_col(fill = "aquamarine4") +
  geom_text(
    aes(n_20s, twenties, label = round(n_20s, 2)),
    nudge_x = -2,
    size = 4
  ) +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those <span style = 'color: aquamarine4'>**20-30**</span>",
  ) +
    theme(plot.title = ggtext::element_markdown()) 

  
# plotting the oldest age group--again, cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
n_over30 <- age$n[31:45]
over_30 <- age$`Over 30`[31:45]

oldest <- data.frame(n_over30, over_30)

  ggplot(oldest, aes(n_over30, over_30)) +
  geom_col(fill = "lemonchiffon4") +
    geom_text(
    aes(n_over30, over_30, label = round(n_over30, 2)),
    nudge_x = -2,
    size = 4
  ) +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those <span style = 'color: lemonchiffon4'>**over 30**</span>",
  ) +
    theme(plot.title = ggtext::element_markdown()) 
```


# Research question 2b: 
## What is the distribution of topics discussed?



### For Final:
#### -Make percentages of each slice more clear in final pie chart (use pie() rather than coord_polar(theta)?)
#### -Figure out what the largest category variable in the pie chart ('indUnk') means.



```{r rq2b, warning= FALSE, echo = FALSE, message = FALSE}

tidy_words$topic <- as.factor(tidy_words$topic)
# levels(tidy_words$topic)


topics <- tidy_words %>% 
  group_by(topic) %>%
  summarise(topic_totals = n()) %>%
  mutate(topic_percent = topic_totals / sum(topic_totals) * 100) %>%
  slice(20, 36, 14, 37, 5, 15) %>%
  ggplot(aes(x = "", y = topic_percent, fill = topic)) +
  geom_col() +
  coord_polar(theta = "y") + 
  labs(title = "Proportion of blogs covering 5 most popular topics", 
       x = "", y = "") +
  theme_minimal()
  

topics

```

# Research question 3:
## Is there a relationship between blog post length and date posted? 

```{r q3, warning= FALSE, echo = FALSE, message = FALSE}
#new df, converting the date column to dates
rq3 <- data
rq3$date <- gsub(",", "/", rq3$date)
#head(rq3)
rq3$date <- gsub("January", "01", rq3$date)
rq3$date <- gsub("February", "02", rq3$date)
rq3$date <- gsub("March", "03", rq3$date)
rq3$date <- gsub("April", "04", rq3$date)
rq3$date <- gsub("May", "05", rq3$date)
rq3$date <- gsub("June", "06", rq3$date)
rq3$date <- gsub("July", "07", rq3$date)
rq3$date <- gsub("August", "08", rq3$date)
rq3$date <- gsub("September", "09", rq3$date)
rq3$date <- gsub("October", "10", rq3$date)
rq3$date <- gsub("November", "11", rq3$date)
rq3$date <- gsub("December", "12", rq3$date)
#head(rq3)
rq3$date <- as.Date(rq3$date, format="%d/%m/%Y")
#head(rq3)

#word counts for each blog post
#"cont of the number of words for every id/text/date combination"
#"summing within id’s"
#NOTE: I changed the alpha just a small amount to make the lines more clear, it all seemed to bleed together a bit and this may clear it up while still showing all of the points! I also changed the title a bit to make it flow a little better, definitely not necessary just a recommendation. 
rq3_plot <- rq3 %>% 
  select(id, text, date, agegroup) %>% 
  group_by(id, text, date, agegroup) %>% 
  unnest_tokens(word, text) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x=date, 
             y=count,
             colour = agegroup)) +
  geom_point(alpha = .2) +
  theme_bw() +
  scale_color_brewer(palette="Blues") +
  scale_y_log10(labels = scales::comma) +
  scale_x_date(date_labels = "%Y", expand = c(0,0)) +
  geom_smooth(method=lm, se=FALSE) +
  labs(
    x = "Date Posted",
    y = "Number of Words",
    title = "Word Count by Blogger per Date Posted",
    color = "Blogger Age"
  ) 

rq3_plot

```


WOW! Great flow of work! I like how you decided to use the data. I think there was a great progression throughout the project showing different aspects of the data in different ways. I think the use of multiple different plots was cool, I really liked the pie chart and how you used it, will be very neat when you have your final version I agreed with your notes to clarify the percentages. I made a few code suggestions for the geom_col to add in the number on the columns to show how many in each (although I was struggling to position it right), but if you can get that to work I think it would be a great addition to really display the data you are showing. I also thought adding in some ggtext to your titles would add to the visual of what you are plotting to connect the different colors of the columns to the age group you are presenting. There were some additoinal code suggestions I made notes on throughout your project but they are simply a suggestion, I think what you have is great just some different things that could add to the overall visual effect. Overall I loved this project I think you guys are on a great track and I cannot wait to see the final output! 
