---
title: "EDLD652 Final"
author: "Diana Dewald, Eliott Doyle"
date: "2/10/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(rio)
library(here)
library(tidytext)
library(RColorBrewer)

#opening and removing about 90% of the rows since the file was too big to push! This code won't work since I got rid of the blogtext.csv file 
#data1 <- import(here("blogtext.csv"))
#data1 <- data1[-(60000:681284),]
#write.csv(data1, "blogtextabbr.csv")

#reading in the new df with fewer observations
data <- import(here("blogtextabbr.csv"))
```

```{r setup2, include=FALSE}
#making all text in 'text' lowercase
data <- data %>% 
  mutate(text = tolower(text))

#View(data)

m_age <- mean(data$age, na.rm = T)  # did not know you can just use 'T', cool!
m_age
max_age <- max(data$age, na.rm = T)
max_age
min_age <- min(data$age, na.rm = T)
min_age

#creating a new column with three groups for age
data <- data %>% 
  mutate(agegroup = cut(age, 
                        breaks =c (-Inf, 19.9, 30, Inf), 
                        labels=c("Under 20","20s","Over 30")))
head(data)
```

# Description of the data
#### The data used for this project are from https://www.kaggle.com/rtatman/blog-authorship-corpus, a text dataset of approximately 681000 blogposts. For the purposes of this project, we will be working with one tenth of the full dataset. In addition to the text of each blog post, some information about each the post (topic, date posted) and blogger (gender, age, astrological sign) is included.

# Research question 1:
## What are the most frequently used words in all writing samples?

```{r rq1, warning= FALSE, echo = FALSE, message = FALSE}
#parsing words from the 'text' column
tidy_words <- data %>% 
  unnest_tokens(word, text)

#removing numbers
tidy_words <- tidy_words[-grep("\\b\\d+\\b", tidy_words$word),]

#removing common/under-informative words
exclu <- tibble(word = c("nbsp","urllink", "lot", "â", "im", "pm", "blog", "stuff", "lol"))
tidy_words <- tidy_words %>% 
  anti_join(exclu, by = "word")
head(tidy_words)

#counting most frequently occurring words, removing stop words from counts
# tidy_words %>% 
#  anti_join(stop_words) %>%
#  count(word, sort = TRUE)

#plot
tidy_words %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% # make y-axis ordered by n
  slice(1:15) %>% # select only the first 15 rows
  ggplot(aes(n, word)) +
  geom_col(fill = "cornflowerblue") +
  scale_x_continuous(expand = c(0,0)) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80"),
    axis.title.x = element_text(face = "bold", size = 12),
    axis.title.y = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold", size =16)
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words across all blog posts",
    caption = "Data abbreviated from https://www.kaggle.com/rtatman/blog-authorship-corpus"
  )

### Cool code used, I had a hard time with string manipulation so glad to see that you are all able to incorporate for the project. I recommend polishing the fonts on axis and title like I added in your code above to make them pop more. Also recommend adding the actual value to time since it looks like it extends beyond the visual? Sorry, didn't add that code in.
```

# Research questions 2a: 
## What are the most frequently used words by age group? Which words are unique to certain age groups?

### For Final:
#### -color columns by words that are unique to each age group, as well as words that are used in only 2 age groups

```{r rq2a, warning= FALSE, echo = FALSE, message = FALSE}
# separate words by valence (positive and negative)?
# separate words by categories (time, place, noun, feeling)?

# I was thinking maybe you could do the same (what you said above) but not sure how to specify categories (can you tell if the word was meant positively or negatively?). Maybe by categories could be better?

# what I tried initially to organize dataframe by age (wasn't ideal for plotting)
# words_age <- tidy_words %>%
#  mutate(under_20 = ifelse(agegroup == "Under 20", 1, 0),
#         "20s" = ifelse(agegroup == "20s", 1, 0),
#         over_30 = ifelse(agegroup == "Over 30", 1, 0)) %>%
#  select(under_20, word)

#also tried group_by(agegroup) in initial plot, this didn't work, nor did facet_wrap(~agegroup)
#I know the code below is not the most elegant way to group by age, so I'm open to suggestions here

# note: I attempted to color bars by uniqueness (see 'for final' not above), but leftjoin() was giving me trouble


# prepping data
tidy <- tidy_words %>% 
  group_by(agegroup) %>%
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% 
  slice(1:15)

age <- tidy %>%
  distinct() %>%
  pivot_wider(names_from = agegroup, values_from = word) 


# plotting younget age group
n_under20 <- age$n[1:15]
under_20 <- age$`Under 20`[1:15]

young <- data.frame(n_under20, under_20)


  ggplot(young, aes(n_under20, under_20)) +
  geom_col(fill = "lightpink4") +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those under 20",
  )


# plotting middle age group--cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
n_20s <- age$n[16:30]
twenties <- age$`20s`[16:30]

middle <- data.frame(n_20s, twenties)
middle$n_20s <- as.double(middle$n_20s)

middle2 <- middle %>% 
  arrange(desc(n_20s)) 

middle2 %>% 
  ggplot(aes(n_20s, twenties)) +
  geom_col(fill = "aquamarine4") +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
      scale_x_continuous(expand = c(0,0)) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those 20-30",
  )

  
# plotting the oldest age group--again, cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
n_over30 <- age$n[31:45]
over_30 <- age$`Over 30`[31:45]

oldest <- data.frame(n_over30, over_30)

  ggplot(oldest, aes(n_over30, over_30)) +
  geom_col(fill = "lemonchiffon4") +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those over 30",
  )
  
# Looking at the three age groups' visuals, the top 5 words are the same across. I don't know that highlighting unique words for each group (with gghighlight) would be good, as opposed to separating by categories as you mention. I also had trouble reordering by most to least. I even created another df in descending order and it wouldn't change the visual. :(
```


# Research question 2b: 
## What is the distribution of topics discussed?



### For Final:
#### -Make percentages of each slice more clear in final pie chart (use pie() rather than coord_polar(theta)?)
#### -Figure out what the largest category variable in the pie chart ('indUnk') means.



```{r rq2b, warning= FALSE, echo = FALSE, message = FALSE}

tidy_words$topic <- as.factor(tidy_words$topic)
# levels(tidy_words$topic)


topics <- tidy_words %>% 
  group_by(topic) %>%
  summarise(topic_totals = n()) %>%
  mutate(topic_percent = topic_totals / sum(topic_totals) * 100) %>%
  slice(20, 36, 14, 37, 5, 15) %>%
  ggplot(aes(x = "", y = topic_percent, fill = topic)) +
  geom_col() +
  coord_polar(theta = "y") + 
  labs(title = "Proportion of blogs covering 5 most popular topics", 
       x = "", y = "") +
  theme_minimal()


topics

# This is super neat, hadn't seen or even considered a pie chart. Great work on getting it this far. Adding in percentages of each would be a big plus to this! I was also wondering what 'indUnk' means.
```

# Research question 3:
## Is there a relationship between blog post length and date posted? 

```{r q3, warning= FALSE, echo = FALSE, message = FALSE}
#new df, converting the date column to dates
rq3 <- data
rq3$date <- gsub(",", "/", rq3$date)
#head(rq3)
rq3$date <- gsub("January", "01", rq3$date)
rq3$date <- gsub("February", "02", rq3$date)
rq3$date <- gsub("March", "03", rq3$date)
rq3$date <- gsub("April", "04", rq3$date)
rq3$date <- gsub("May", "05", rq3$date)
rq3$date <- gsub("June", "06", rq3$date)
rq3$date <- gsub("July", "07", rq3$date)
rq3$date <- gsub("August", "08", rq3$date)
rq3$date <- gsub("September", "09", rq3$date)
rq3$date <- gsub("October", "10", rq3$date)
rq3$date <- gsub("November", "11", rq3$date)
rq3$date <- gsub("December", "12", rq3$date)
#head(rq3)
rq3$date <- as.Date(rq3$date, format="%d/%m/%Y")
#head(rq3)

#word counts for each blog post
#"cont of the number of words for every id/text/date combination"
#"summing within id’s"
rq3_plot <- rq3 %>% 
  select(id, text, date, agegroup) %>% 
  group_by(id, text, date, agegroup) %>% 
  unnest_tokens(word, text) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x=date, 
             y=count,
             colour = agegroup)) +
  geom_point(alpha = .5) +
  theme_bw() +
  scale_color_brewer(palette="Blues") +
  scale_y_log10(labels = scales::comma) +
  scale_x_date(date_labels = "%Y", expand = c(0,0)) +
  geom_smooth(method=lm, se=FALSE) +
  labs(
    x = "Date Posted",
    y = "Number of Words",
    title = "Word Count by Blogger by Date",
    color = "Blogger Age"
  ) 

rq3_plot

# This plot is super cool and gets the message across! Love it. I think one thing to note is a comment somewhere to be aware of the y-axis and the scale there. Nice work!
```



