---
title: "EDLD652 Final"
author: "Diana Dewald, Eliott Doyle"
date: "2/10/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(rio)
library(here)
library(tidytext)
library(RColorBrewer)

#opening and removing about 90% of the rows since the file was too big to push! This code won't work since I got rid of the blogtext.csv file 
#data1 <- import(here("blogtext.csv"))
#data1 <- data1[-(60000:681284),]
#write.csv(data1, "blogtextabbr.csv")

#reading in the new df with fewer observations
data <- import(here("blogtextabbr.csv"))
```

```{r setup2, include=FALSE}
#making all text in 'text' lowercase
data <- data %>% 
  mutate(text = tolower(text))

#View(data)

m_age <- mean(data$age, na.rm = T)
m_age
max_age <- max(data$age, na.rm = T)
max_age
min_age <- min(data$age, na.rm = T)
min_age

#creating a new column with three groups for age
data <- data %>% 
  mutate(agegroup = cut(age, 
                        breaks =c (-Inf, 19.9, 30, Inf), 
                        labels=c("Under 20","20s","Over 30")))
head(data)
```

# Description of the data
#### The data used for this project are from https://www.kaggle.com/rtatman/blog-authorship-corpus, a text dataset of approximately 681000 blogposts. For the purposes of this project, we will be working with one tenth of the full dataset. In addition to the text of each blog post, some information about each the post (topic, date posted) and blogger (gender, age, astrological sign) is included.

# Research question 1:
## What are the most frequently used words in all writing samples?

```{r rq1, warning= FALSE, echo = FALSE, message = FALSE}
#parsing words from the 'text' column
tidy_words <- data %>% 
  unnest_tokens(word, text)

#removing numbers
tidy_words <- tidy_words[-grep("\\b\\d+\\b", tidy_words$word),]

#removing common/under-informative words
exclu <- tibble(word = c("nbsp","urllink", "lot", "â", "im", "pm", "blog", "stuff", "lol"))
tidy_words <- tidy_words %>% 
  anti_join(exclu, by = "word")
head(tidy_words)

#counting most frequently occurring words, removing stop words from counts
# tidy_words %>% 
#  anti_join(stop_words) %>%
#  count(word, sort = TRUE)

#plot
tidy_words %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% # make y-axis ordered by n
  slice(1:15) %>% # select only the first 15 rows
  ggplot(aes(n, word)) +
  geom_col(fill = "cornflowerblue") +
  scale_x_continuous(expand = c(0,0)) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words across all blog posts",
    caption = "Data abbreviated from https://www.kaggle.com/rtatman/blog-authorship-corpus"
  )


## This is really interesting - I wonder what your takeaway is here. This gives me the sense that people are v philosophical and musing on the nature of time and people and love and life. My main suggestion for this plot was going to be to try and show some of the differences in word use by different groups, which it looks like you're already doing in your next section, so excellent job of beating me to it!
```

# Research questions 2a: 
## What are the most frequently used words by age group? Which words are unique to certain age groups?

### For Final:
#### -color columns by words that are unique to each age group, as well as words that are used in only 2 age groups

```{r rq2a, warning= FALSE, echo = FALSE, message = FALSE}
# separate words by valence (positive and negative)?
# separate words by categories (time, place, noun, feeling)?


# what I tried initially to organize dataframe by age (wasn't ideal for plotting)
# words_age <- tidy_words %>%
#  mutate(under_20 = ifelse(agegroup == "Under 20", 1, 0),
#         "20s" = ifelse(agegroup == "20s", 1, 0),
#         over_30 = ifelse(agegroup == "Over 30", 1, 0)) %>%
#  select(under_20, word)

#also tried group_by(agegroup) in initial plot, this didn't work, nor did facet_wrap(~agegroup)
#I know the code below is not the most elegant way to group by age, so I'm open to suggestions here

# note: I attempted to color bars by uniqueness (see 'for final' not above), but leftjoin() was giving me trouble


# prepping data
tidy <- tidy_words %>% 
  group_by(agegroup) %>%
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% 
  slice(1:15)

age <- tidy %>%
  distinct() %>%
  pivot_wider(names_from = agegroup, values_from = word) 


# plotting younget age group
n_under20 <- age$n[1:15]
under_20 <- age$`Under 20`[1:15]

young <- data.frame(n_under20, under_20)


young_plot <-   ggplot(young, aes(n_under20, under_20)) +
  geom_col(fill = "lightpink4") +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those under 20",
  )

young_plot

# plotting middle age group--cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
  
  
  ## wow i could not get the reordering to work either after 20 minutes of audibly arguing with my computer...
  
n_20s <- age$n[16:30]
twenties <- age$`20s`[16:30]

middle <- data.frame(n_20s, twenties)

mid_plot <-   ggplot(middle, aes(n_20s, twenties)) +
  geom_col(fill = "aquamarine4") +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those 20-30",
  )

mid_plot

# plotting the oldest age group--again, cannot get the re-ordering to cooperate here for the y-axis, tried several solutions to no avail, any edits welcome
n_over30 <- age$n[31:45]
over_30 <- age$`Over 30`[31:45]

oldest <- data.frame(n_over30, over_30)

oldest_plot <-   ggplot(oldest, aes(n_over30, over_30)) +
  geom_col(fill = "lemonchiffon4") +
  theme_minimal() +
  scale_x_continuous(expand = c(0,0)) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(color = "gray80")
  ) +
  labs(
    x = "Word Frequency",
    y = "Word",
    title = "Top 15 most frequently occurring words for those over 30",
  )
oldest_plot



## Here's a workaround for facet-gridding the plots that makes it a bit more straightforward to compare the plots. 
#If you go with this method I'd recommend some retitling, and some highlighting or formatting to allow for comparing words across ages. For example, I might suggest using the same color for each word across groups, and unique colors for words that are only in one age group (now that I write that I think my idea is 1. a color for words in common across all groups 2. a color for words only common across 2 groups and 3. unique colors for each group for the words only they use)
#Something like that, idk, but the idea would be to make it easier to emphasize what you think is most interesting about the combined plots
library(ggpubr)
ggarrange(young_plot, mid_plot, oldest_plot,
                    ncol = 2, nrow = 2)
```


# Research question 2b: 
## What is the distribution of topics discussed?



### For Final:
#### -Make percentages of each slice more clear in final pie chart (use pie() rather than coord_polar(theta)?)
#### -Figure out what the largest category variable in the pie chart ('indUnk') means.



```{r rq2b, warning= FALSE, echo = FALSE, message = FALSE}

tidy_words$topic <- as.factor(tidy_words$topic)
# levels(tidy_words$topic)


topics <- tidy_words %>% 
  group_by(topic) %>%
  summarise(topic_totals = n()) %>%
  mutate(topic_percent = topic_totals / sum(topic_totals) * 100) %>%
  slice(20, 36, 14, 37, 5, 15) %>%
  ggplot(aes(x = "", y = topic_percent, fill = topic)) +
  geom_col() +
  coord_polar(theta = "y") + 
  labs(title = "Proportion of blogs covering 5 most popular topics", 
       x = "", y = "") +
  theme_minimal()
  

topics


# For this plot it might help to expand what the indUnk topic means - I'm not sure I could guess the topic and it appears to be the most common topic. I know pie charts are not supposed to be a thing...but I kinda like how it conveys the information here. Another recommendation I might make that is *very subjective* is changing the color palette, maybe it's bc I'm enjoying the dark and cloudy day but these colors are very bright
```

# Research question 3:
## Is there a relationship between blog post length and date posted? 

```{r q3, warning= FALSE, echo = FALSE, message = FALSE}
#new df, converting the date column to dates
rq3 <- data
rq3$date <- gsub(",", "/", rq3$date)
#head(rq3)
rq3$date <- gsub("January", "01", rq3$date)
rq3$date <- gsub("February", "02", rq3$date)
rq3$date <- gsub("March", "03", rq3$date)
rq3$date <- gsub("April", "04", rq3$date)
rq3$date <- gsub("May", "05", rq3$date)
rq3$date <- gsub("June", "06", rq3$date)
rq3$date <- gsub("July", "07", rq3$date)
rq3$date <- gsub("August", "08", rq3$date)
rq3$date <- gsub("September", "09", rq3$date)
rq3$date <- gsub("October", "10", rq3$date)
rq3$date <- gsub("November", "11", rq3$date)
rq3$date <- gsub("December", "12", rq3$date)
#head(rq3)
rq3$date <- as.Date(rq3$date, format="%d/%m/%Y")
#head(rq3)

#word counts for each blog post
#"cont of the number of words for every id/text/date combination"
#"summing within id’s"
rq3_plot <- rq3 %>% 
  select(id, text, date, agegroup) %>% 
  group_by(id, text, date, agegroup) %>% 
  unnest_tokens(word, text) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x=date, 
             y=count,
             colour = agegroup)) +
  geom_point(alpha = .5) +
  theme_bw() +
  scale_color_brewer(palette="Blues") +
  scale_y_log10(labels = scales::comma) +
  scale_x_date(date_labels = "%Y", expand = c(0,0)) +
  geom_smooth(method=lm, se=FALSE) +
  labs(
    x = "Date Posted",
    y = "Number of Words",
    title = "Word Count by Blogger by Date",
    color = "Blogger Age"
  ) 

rq3_plot


# This is really really cool. If there's any way to get data after 2004 I'd be suuuuper interested in seeing is the explosion of word count continues into the latter aughts. My prediction is that things peaked in the earlyish 2010's and we've moved to a pithier blog style since. Also, who are these people making one word blog posts?! So many questions come up with this plot, which underscores the fact that this is a really cool idea. If I were to suggest something beyond adding more data that you might not have access to, it would be to choose a color palette that has more contrast between the groups


```



